### 2010-01-25 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
draft for deliverable 5.1.2 (M24)
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@471 d8b0c59b-6544-0410-94d9-83a5658a058e

\newcommand\deliverableNumber{D5.2.1}
\newcommand\deliverableTitle{Prototype Semantic Catalogue for Screen-Flow Gadgets and Back-End Services}
\newcommand\deliverableTitleShort{Prototype Semantic Catalogue}
\newcommand\workpackageNumber{5}
\newcommand\workpackageTitle{Semantic catalogue of screen-flow resources and back-end Web Services}
\newcommand\authorOne{Ismael Rivera, NUIG}
\newcommand\authorTwo{Author 2}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}

@@

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 20/02/2009}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                         \textcolor{white}{\textbf{Date}} &
                         \textcolor{white}{\textbf{Author (Partner)}} &
                         \textcolor{white}{\textbf{Change description}}\\ \hline
1.0 & 20/02/2009 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
\end{tabular}
\end{small}

@@

\begin{small}
\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{Status} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable is the technical documentation for the prototype developed as part of the WP5. It describes the catalogue's architecture, data interchange formats, APIs and examples of how to interact with it.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\
  \hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495086} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, semantic catalogue, gadget catalogue, RDF store}\\ \hline
\end{tabular}
\end{small}

@@

\clearpage
\section{Introduction}
\label{sec:introduction}
This section starts establishing the goal and scope of the present document, shows how it is structured and details the relation to others documents and work packages.

\subsection{Goal and Scope}
This is an introductory manual for developers who want to adopt and use the FAST semantic catalogue. It explains the main functionalities implemented so far, an overview of its architecture and a detailed \emph{Application Programming Interface} or API of the complete set of the operations offered through a REST service.

\subsection{Structure of the document}
The deliverable presents both the external and internal architecture in Section \ref{sec:architecture}, then in Section \ref{sec:api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on, and the document ends by stating some conclusions and future directions of the semantic catalogue in Section \ref{sec:conclusions}.

\subsection{Overall approach for WP5 and relation to other Work Packages and Deliverables}
The main objective of the Work Package 5, as specified in \cite{DoW}, includes the development of the FAST semantic catalogue, considered the back-end of the FAST platform, which will store and index any kind of resources such as screen-flows, screens, back-end Web services descriptions, locate appropriate gadgets, resources and Web services for users, storing and indexing user profiles, and resolve mediation problems between gadgets, back-end services and the different ontologies stored.

The deliverables stated in \cite{DoW} for this WP are:
\begin{description}
	\item[User manual of the FAST Catalogue] This is the user manual of the catalogue, but is out of scope in this iteration of the project.
	\item[Prototype semantic catalogue for screen-flow gadgets and back-end services] This deliverable is a software prototype of the semantic catalogue for experimentation purposes. Apart from the software, it will be created technical documentation and manuals for developers as the present document.
\end{description}

@@

\clearpage
\section{Architecture}
\label{sec:architecture}
This section details the internal catalogue's architecture and which external components will interact with it. Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure \ref{fig:architecture}).

@@

\clearpage
\section{RESTful Catalogue API}
\label{sec:api}
This section provides a high-level overview of the Catalogue API. It describes the calls or operations supported, specific parameters and responses for each operation, supported interchange formats and some examples to facilitate the understanding of the API.

\subsection{JSON Interchange Format}
The format supported by the Catalogue API for all the request and responses is JSON\footnote{http://www.json.org/}. JSON is a lightweight data interchange format whose simplicity has resulted in widespread use among web developers, easy to read and write and able to using any programming language because its structures map directly to data structures used in most programming languages.

@@

@@

\subsubsection{Internationalisation I18n}
In order to offers an adaptable solution to various languages and regions without major engineering changes, internationalisation is considered from an early stage in the catalogue development. Underlying representation technologies used to develop the catalogue (RDF and JSON) allow to implement this feature. This feature is implemented by the addition of a language tag to every 'string' desired. The specification of this language tag is composed by the language code and the country code, following the ISO 639\footnote{http://ftp.ics.uci.edu/pub/ietf/http/related/iso639.txt} for languages and the ISO 3166\footnote{http://userpage.chemie.fu-berlin.de/diverse/doc/ISO\_3166.html} for countries. The following example illustrates how to use it properly in both formats.

@@

\subsection{API Calls}
This section contains detailed descriptions of the interface provided by the catalogue via REST services, detailing request parameters, response elements, any special errors and examples of requests and responses. The URL format is also specified for each operation, where 'catalogueURL' has to be replaced for the real URL the service is installed (e.g. forge.morfeo-project.org/catalogue).

\subsubsection{CRUD operations for Screens}
A screen is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screens. These features are provided following the REST philosophy, using the HTTP methods. Two concepts have to be defined: \emph{collection}, as a set of screens which access URI is http://catalogueURL/screens/ where catalogueURL is the specific URL where the catalogue server is installed, and \emph{member} of the collection, in other words, the screen itself, which access URI is http://catalogueURL/screens/<id> where the <id> has to be replaced by the identifier of a screen. The details of the operations and which HTTP verb has to be used can be found in the Table \ref{tab:crud_screens}.

@@

\subsubsection{CRUD operations for Screen-Flows}
A screenflow is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screenflows. These features are provided following the REST philosophy, using the HTTP methods. As in the previous section the concepts \emph{collection} and \emph{member} are used including their URLs. Table \ref{tab:crud_screenflows} details the operations offered for screen-flows.

@@

\subsubsection{Find}
The find operation search inside the catalogue for any screen which could be somehow related to the gadget the user is creating. It provides a recommended set of screens depending on the domain context, the canvas, etc. The specific URL to access this operation is http://catalogueURL/find using HTTP POST method. From now on, the method 'find' only considers the domain context, being a list of tags and the pre/postconditions of the screens. It will try to satisfy all the unsatisfied preconditions of a given list of screens also known as canvas. The results will be all the screens stored in the catalogue which fulfil some of these preconditions, and are tagged with the tags specified in the domain context. The request parameters are shown in Table \ref{tab:find_request} and the response parameters are shown in Table \ref{tab:find_response}.

@@

\subsubsection{Check}
This operation verifies the state of certain list of screens depending on a specific criterion. The only criterion accepted in this stage of the prototype is 'reachability'. A screen will be reachable if it all its preconditions are satisfied by postconditions of the screens the canvas contain. The specific URL to access this operation is http://catalogueURL/check using HTTP POST method.

@@

\subsubsection{GetMetadata}
Although the information about a resource can be obtained by its specific retrieval operation supported by the CRUD interface, sometimes is needed to retrieve information which would imply a request per resource, hence this operation allow to get the metadata of a list of resources in only one request. This prototype supports this operation for screenflows and screens. The specific URL to access this operation is http://catalogueURL/getmetadata using HTTP POST method. Table \ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table \ref{tab:getmetadata_response} details the different parameters may contain a response.

@@

\subsubsection{FindAndCheck}
Not supported yet.

\subsection{API Error Codes}
There are two types of error codes, client and server.
\begin{itemize}
	\item Client error codes are generally caused by the client and might be an invalid domain or an invalid request parameter. These errors are accompanied by a 4xx HTTP response code. For example: ResourceNotFound.
	\item Server error codes are generally caused by a server-side issue and should be reported. These errors are accompanied by a 5xx HTTP response code. For example: ServerUnavailable.
\end{itemize}

@@

\clearpage
\section{Conclusions and Future Work}
\label{sec:conclusions}
As a result of the work done for the back-end catalogue, any one who read this document should be able to understand the capabilities implemented and how to use them. Summarising, this prototype allows the storage of screens and the composition of screen-flows. Future work will involve the storage of the internal components of a screen, such as connectors, filters, forms and back-end services. However, the most interesting work will be done in the optimization of the resources recommendation feature, in other words, it will be investigate new mechanisms to improve the results of the find such as weighting of answers, planning, etc. Last of all, more complicated queries will be taken into account, improving the matching algorithm between pre and postconditions.

\clearpage
\bibliographystyle{apalike}
\bibliography{fast_bib}

\clearpage
\doublespacing
\section*{Appendix A (Lists of Tables and Figures)}
\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}

@@

>>>>

\newcommand\deliverableNumber{D5.1.2}
\newcommand\deliverableTitle{User Manual of the FAST Catalogue}
\newcommand\deliverableTitleShort{Catalogue User Manual}
\newcommand\workpackageNumber{5}
\newcommand\workpackageTitle{Semantic catalogue of screen-flow resources and back-end Web Services}
\newcommand\authorOne{Ismael Rivera, NUIG}
\newcommand\authorTwo{Author 2}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

@@

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 25/01/2010}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                         \textcolor{white}{\textbf{Date}} &
                         \textcolor{white}{\textbf{Author (Partner)}} &
                         \textcolor{white}{\textbf{Change description}}\\ \hline
1.0 & 25/01/2010 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
\end{tabular}
\end{small}

@@

\begin{small}
\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{Status} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable is the technical documentation for the prototype developed as part of the WP5. It describes the catalogue's architecture, data interchange formats, APIs and examples of how to interact with it.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\
  \hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{NUIG} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495338} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, semantic catalogue, gadget catalogue, RDF store}\\ \hline
\end{tabular}
\end{small}

@@

\clearpage
\section{Introduction} % (fold)
\label{sec:introduction}
This section starts establishing the goal and scope of the present document, shows how it is structured and details the relation to others documents and work packages.

\subsection{Goal and Scope} % (fold)
\label{sub:goal_and_scope}

This is an introductory manual for developers who want to adopt and use the FAST semantic catalogue. It explains the main functionalities implemented so far, an overview of its architecture and a detailed \emph{Application Programming Interface} or API of the complete set of the operations offered through a REST service.

% subsection goal_and_scope (end)

\subsection{Structure of the document} % (fold)
\label{sub:structure_of_the_document}

The deliverable presents both the external and internal architecture in Section \ref{sec:architecture_overview}, then in Section \ref{sec:restful_catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.

% subsection goal_and_scope (end)

\subsection{Overall approach for WP5 and relation to other Work Packages and Deliverables} % (fold)
\label{sub:overall_approach_for_wp5}

The main objective of the Work Package 5, as specified in \cite{DoW}, includes the development of the FAST semantic catalogue, considered the back-end of the FAST platform, which will store and index any kind of resources such as screen-flows, screens, back-end Web services descriptions, locate appropriate gadgets, resources and Web services for users, storing and indexing user profiles, and resolve mediation problems between gadgets, back-end services and the different ontologies stored.

The deliverables stated in \cite{DoW} for this WP are:
\begin{description}
	\item[User manual of the FAST Catalogue] This is the user manual of the catalogue. Since the semantic catalogue will not provide a user graphical interface by itself: it relies on the GVS; this document will be oriented to developers, explaining technical aspects of the catalogue, its API and several examples to ease its use.
	\item[Prototype semantic catalogue for screen-flow gadgets and back-end services] This deliverable is a software prototype of the semantic catalogue for experimentation purposes.
\end{description}

@@

% subsection overall_approach_for_wp5 (end)

\subsection{Changes from D5.2.1} % (fold)
\label{sub:changes}

This should be the first version of this deliverable since in M12 the deliverable 5.1 was not scheduled. Nevertheless, last year an extra document was written as an appendix to the FAST Catalogue in order to detail and clarify technical decisions and to have a technical reference manual. Hence, the deliverable D5.2.1 will evolve to become the D5.1.2, so it will be no longer needed by itself.

There are several new sections and major changes which need to be mentioned:

\begin{itemize}
	\item Section \ref{sec:installation_guide}. Installation Guide describes a set of steps to be followed in order to install and execute the FAST Catalogue. Also some configuration instructions are specified.
	\item Section \ref{sub:content_negotiation}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
	\item Section \ref{sec:restful_catalogue_api}. RESTful Catalogue API has been intensively extended and modified in order to  reflect all the new functionality and how every request or response is constructed.
	\item Several appendixes have been added with the JSON structure of every building block the Catalogue can deal with.
\end{itemize}

% subsection changes (end)

% section introduction (end)

\clearpage
\section{Installation Guide} % (fold)
\label{sec:installation_guide}

This section  provides instructions on how to manually install and configure the FAST Catalogue. The first part of this guide presents some requirements to be considered before the installation, then it gives broad general instructions, and the last part contains a more detailed installation notes for specific configurations.

\subsection{System Requirements} % (fold)
\label{sub:system_requirements}

In addition to the software itself, a standard FAST Catalogue installation has the following requirements:
\begin{itemize}
  \item Java\footnote{http://java.sun.com/} (version 6 or later) is required to run the software.
  \item A Java Servlet Container that supports Java Servlet API 2.4 and Java Server Pages (JSP) 2.0, or newer. We recommend using a recent, stable version of Apache Tomcat\footnote{http://tomcat.apache.org/}. At the time of writing, this is either version 5.5.x or 6.x.
\end{itemize}

In addition, there are various optional dependencies which are required if you want to use certain advanced features (see Section \ref{sub:installation_instructions}).

% subsection system_requirements (end)

\subsection{Obtaining the FAST catalogue} % (fold)
\label{sub:obtaining_the_fast_catalogue}

It is recommended to have a Subversion client installed before you download the code (although you can theoretically download files without Subversion, this would mean tediously downloading each individual file manually). The recommended software is the official Subversion client, available from the Subversion project page\footnote{http://subversion.tigris.org/}. Note that this client uses a command-line interface, which the instructions below use. Alternatively, you can get subversioning software with a graphical user interface such as TortoiseSVN.

To download from the latest release (recommended), enter the following command from the command-line in the directory you wish to download to:
\begin{verbatim}
svn checkout https://svn.forge.morfeo-project.org/fast-fp7project/trunk/catalogue_service
\end{verbatim}

% subsection obtaining_the_fast_catalogue (end)

\subsection{Installation Instructions} % (fold)
\label{sub:installation_instructions}

Once you have got the source code, first step should be addressed is its compilation. You may manually compile the source yourself, however the FAST Catalogue comes along with a script to facilitate this task. This script has been made using a Java build tool called Ant \footnote{http://ant.apache.org/}. Hence, we recommended to get and install such a tool in order to follow the instructions below.

Now you can compile, package and run the application via:
\begin{verbatim}
ant clean
ant compile
ant dist
\end{verbatim}

Or using a task which gather the previous three tasks:
\begin{verbatim}
ant main
\end{verbatim}

After executing with no errors this task, a WAR file should have been created in the /dist directory. A WAR file, which stands for  Web Application Archive, is just a JAR file used to distribute a collection of JavaServer Pages, servlets, Java classes, XML files, tag libraries and static Web pages (HTML and related files) that together constitute a Web application, in this case, the FAST Catalogue.

Then, last step is to deply the WAR file into the servlet container. If the server chosen was Apache Tomcat, the procedure for deploying a Web application is:
\begin{enumerate}
	\item Stop Tomcat.
	\item Delete existing deployment. If you have previously deployed "foo.war" in TOMCAT\_HOME/webapps, then it has been unpacked into webapps/foo/... You must delete this directory and all its contents. On Unix, this can be done with
rm -r \$TOMCAT\_HOME/webapps/foo
	\item Copy WAR file to TOMCAT\_HOME/webapps/.
	\item Start Tomcat.
\end{enumerate}

Now you have successfully built the FAST catalogue, however we highly recommend you to read Section \ref{sub:configuration} and tune your FAST Catalogue instance.

% subsection installation_instructions (end)

\subsection{Configuration} % (fold)
\label{sub:configuration}

The following section covers the configuration options you may have to set up before start using the Catalogue. The FAST Catalogue relies on a Sesame repository for the persistent storage. Sesame repositories can be local or remote. Remote does not mean, it has to be installed in a different machine, but in a servlet container such as Apache Tomcat. The advantage of remote repositories is the possibility of accessing it via a SPARQL endpoint. These configuration aspect can be found in the file \emph{repository.properties} in the root directory where the Catalogue has been deployed.

Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed. If you are testing the application in your machine, it may be something like \url{http://localhost:8080/FASTCatalogue}.

Local repositories just need the parameter \emph{storageDir} pointing at a local directory, which has write permission in order to allow Sesame to create files in that directory or to delete any file in the directory. For instance:

\begin{verbatim}
storageDir=D:\\FAST\\catalogue\\repository
\end{verbatim}

Nevertheless, setting up a remote repository requires a few more steps. First the OpenRDF Sesame Server needs to be installed. The installation guide of this server is detailed in the Chapter 5 and 6 of the User Guide for Sesame 2.2 \cite{sesame2.2}. Once the Sesame server is deployed, and a repository has been created, the parameters sesameServer and repositoryID will take the values of the Sesame server URL and the repository identifier. For instance:

\begin{verbatim}
sesameServer=http://localhost:8880/openrdf-sesame
repositoryID=c2
\end{verbatim}

% subsection installation_instructions (end)

\subsection{Hello World!} % (fold)
\label{sub:hello_world}

It is recommended to check if the FAST Catalogue has been successfully installed. To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of resource such as Screen. The URL would be similar to \url{http://localhost:8080/FASTCatalogue/screens}. Depending on the variant chosen, the response may be different. For JSON, the response should be \verb|[]|, but in HTML the message \verb|No screens found.| should be shown.

% subsection hello_world (end)

% section installation_guide (end)

\clearpage
\section{Architecture Overview} % (fold)
\label{sec:architecture_overview}

This section presents a technical overview of the internal catalogue's architecture and which external components will interact with it. Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure \ref{fig:architecture}).

@@

% section architecture_overview (end)

\clearpage
\section{RESTful Catalogue API} % (fold)
\label{sec:restful_catalogue_api}

This section provides a high-level overview of the Catalogue API. It describes the calls or operations supported, specific parameters and responses for each operation, supported interchange formats and some examples to facilitate the understanding of the API.

\subsection{JSON Interchange Format} % (fold)
\label{sec:json_interchange_format}

The format supported by the Catalogue API for all the request and responses is JSON\footnote{http://www.json.org/}. JSON is a lightweight data interchange format whose simplicity has resulted in widespread use among web developers, easy to read and write and able to using any programming language because its structures map directly to data structures used in most programming languages.

@@

% subsection json_interchange_format (end)

@@

\subsubsection{Internationalisation I18n} % (fold)

In order to offers an adaptable solution to various languages and regions without major engineering changes, internationalisation is considered from an early stage in the catalogue development. Underlying representation technologies used to develop the catalogue (RDF and JSON) allow to implement this feature. This feature is implemented by the addition of a language tag to every 'string' desired. The specification of this language tag is composed by the language code and the country code, following the ISO 639\footnote{http://ftp.ics.uci.edu/pub/ietf/http/related/iso639.txt} for languages and the ISO 3166\footnote{http://userpage.chemie.fu-berlin.de/diverse/doc/ISO\_3166.html} for countries. The following example illustrates how to use it properly in both formats.

@@

\subsection{Content Negotiation} % (fold)
\label{sub:content_negotiation}

Before any major details about how content negotiation is addressed by the catalogue, a few concepts need to be understood. First of all, say a URI is not a file name. It is common to see a URI as the location for a file, but a URI or Universal Resource Identifier is not a URL: it is not a file name or location, it is an identifier (or a reference) to a resource. Then by dereferencing a URI the actual resource can be retrieved. Let's see a request to retrieve the resource "Screen 123" is sent to the server. The resource is just a piece of information, however the server will answer an HTML document with a text describing the screen or an image representing it. On the Web, these equivalent representations of a resource are called variants, and content negotiation is the mechanism used to determine which of the representations of most appropiate for a given request.

\begin{figure}[htb]
\label{fig:ContentNegotiation}
\begin{center}
	\includegraphics[width=10cm]{images/content_negotiation}
	\caption{Content Negotiation}
\end{center}
\end{figure}

In summary, the basic idea of content negotiation, as stated in the HTTP 1.1 specification \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.

In fact, what the catalogue really does is "format negotiation" since the alternate representationos are just based on the selection of the media type, through the Accept header, but does not consider different languages or encoding types. The formats supported are JSON, RDF+XML and HTML.

Lastly, content negotiation needs to identify which player is going to take the lead on it. There are two kinds of content negotiation which are possible in HTTP: server-driven and agent-driven negotiation. The approach followed by the catalogue is agent-driven negotiation, hence selecting a specific representation for a resource is responsibility of the user agent. If none is specified, by default, the server will choose the JSON representation.

% subsection content_negotiation (end)

\subsection{API Calls} % (fold)
\label{sub:api_calls}

This section contains detailed descriptions of the interface provided by the catalogue via REST services, detailing request parameters, response elements, any special errors and examples of requests and responses. The URL format is also specified for each operation, where 'catalogueURL' has to be replaced for the real URL the service is installed (e.g. http://demo.fast.morfeo-project.org/catalogue).

\subsubsection{CRUD operations for Screens} % (fold)

A screen is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screens. These features are provided following the REST philosophy, using the HTTP methods. Two concepts have to be defined: \emph{collection}, as a set of screens which access URI is http://catalogueURL/screens/ where catalogueURL is the specific URL where the catalogue server is installed, and \emph{member} of the collection, in other words, the screen itself, which access URI is http://catalogueURL/screens/<id> where the <id> has to be replaced by the identifier of a screen. The details of the operations and which HTTP verb has to be used can be found in the Table \ref{tab:crud_screens}.

@@

\subsubsection{CRUD operations for Screen-Flows} % (fold)

A screenflow is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screenflows. These features are provided following the REST philosophy, using the HTTP methods. As in the previous section the concepts \emph{collection} and \emph{member} are used including their URLs. Table \ref{tab:crud_screenflows} details the operations offered for screen-flows.

@@

\subsubsection{Find} % (fold)

The find operation search inside the catalogue for any screen which could be somehow related to the gadget the user is creating. It provides a recommended set of screens depending on the domain context, the canvas, etc. The specific URL to access this operation is http://catalogueURL/find using HTTP POST method. From now on, the method 'find' only considers the domain context, being a list of tags and the pre/postconditions of the screens. It will try to satisfy all the unsatisfied preconditions of a given list of screens also known as canvas. The results will be all the screens stored in the catalogue which fulfil some of these preconditions, and are tagged with the tags specified in the domain context. The request parameters are shown in Table \ref{tab:find_request} and the response parameters are shown in Table \ref{tab:find_response}.

@@

\subsubsection{Check} % (fold)

This operation verifies the state of certain list of screens depending on a specific criterion. The only criterion accepted in this stage of the prototype is 'reachability'. A screen will be reachable if it all its preconditions are satisfied by postconditions of the screens the canvas contain. The specific URL to access this operation is http://catalogueURL/check using HTTP POST method.

@@

\subsubsection{GetMetadata} % (fold)

Although the information about a resource can be obtained by its specific retrieval operation supported by the CRUD interface, sometimes is needed to retrieve information which would imply a request per resource, hence this operation allow to get the metadata of a list of resources in only one request. This prototype supports this operation for screenflows and screens. The specific URL to access this operation is http://catalogueURL/getmetadata using HTTP POST method. Table \ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table \ref{tab:getmetadata_response} details the different parameters may contain a response.

@@

\subsubsection{Screen Find \& Check} % (fold)

Request
\begin{verbatim}
{
  "canvas": [
    { "uri": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860786707" }
  ],
  "elements": [
  ],
  "domainContext": {
    "tags": [],
    "user": null,
  },
  "criterion": "reachability"
}
\end{verbatim}

Response
\begin{verbatim}
{
  "canvas": [{
    "preconditions": [[{
      "id": "sc",
      "label": {"en-gb": "A search criteria"},
      "pattern": "?F http://www.w3.org/1999/02/22-rdf-syntax-ns#type http://aws.amazon.com/AWSECommerceService#SearchCriteria",
      "satisfied": false,
      "positive": true
    }]],
    "reachability": false,
    "uri": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860786707"
  }],
  "elements": [{
    "preconditions": [],
    "reachability": true,
    "uri": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860530032"
  }]
}
\end{verbatim}

% subsection api_calls (end)

\subsection{Planning}
\label{sub:planning}

URL: <CATALOGUE\_URL>/planner
Request
\begin{verbatim}
{
  "goal": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860720035",
  "canvas": []
}
\end{verbatim}

\begin{verbatim}
Response
[
  [
    "http://purl.oclc.org/fast/ontology/gadget#Screen1257866875435",
    "http://purl.oclc.org/fast/ontology/gadget#Screen1257865645621"
  ],
  [
    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860530032",
    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860786707",
    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860772723"
  ]
]
\end{verbatim}

% subsection planning (end)

\subsection{API Error Codes} % (fold)
There are two types of error codes, client and server.
\begin{itemize}
	\item Client error codes are generally caused by the client and might be an invalid domain or an invalid request parameter. These errors are accompanied by a 4xx HTTP response code. For example: ResourceNotFound.
	\item Server error codes are generally caused by a server-side issue and should be reported. These errors are accompanied by a 5xx HTTP response code. For example: ServerUnavailable.
\end{itemize}

@@

\clearpage
\bibliographystyle{apalike}
\bibliography{fast_d5_1_2}

\clearpage
\doublespacing
\section*{Appendix A. Lists of Tables and Figures}
\addcontentsline{toc}{section}{Appendix A. Lists of Tables and Figures}

@@

\clearpage
\section*{Appendix B. Generic Building Block JSON Structure}
\addcontentsline{toc}{section}{Appendix B. Generic Building Block JSON Structure}

\begin{verbatim}
{
  "code": "http://url.com/.../code.js",
  "creationDate": "2009-04-20T17:00:00+0100",
  "creator": "http://www.fabio.es/",
  "description": {"en-gb": "This is a description of the building block"},
  "tags": {
    "tags": [
      "ebay",
      "amazon",
    ],
    "user": <String> // not defined yet
  },
  "homepage": "http://www.homepage.ie/",
  "icon": "http://url.com/images/icon.png",
  "id": "Id of the resource",
  "label": {"en-gb": "Label or title of the building block"},
  "name": "Amazon Item to Ebay Filter",
  "rights": "http://creativecommons.org/",
  "screenshot": "http://www.deri.ie/eBayList-screenshot.jpg",
  "type": "type of the building block", // screenflow, screen, form, operator y resource
  "uri": "http://localhost:8080/catalogue/screens/525",
  "version": "1.0"
}
\end{verbatim}

\clearpage
\section*{Appendix C. Screenflow JSON Structure}
\addcontentsline{toc}{section}{Appendix C. Screenflow JSON Structure}

\begin{verbatim}
{
  <GENERIC BUILDING BLOCK JSON>,
  "contains": [
    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860772723",
    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860720035",
  ],
}
\end{verbatim}

\clearpage
\section*{Appendix D. Screen JSON Structure}
\addcontentsline{toc}{section}{Appendix D. Screen JSON Structure}

\begin{verbatim}
{
  <GENERIC BUILDING BLOCK JSON>,
  "postconditions": [[{
    "id": "url",
    "label": {"en-gb": "Purchase URL"},
    "pattern": "?P
                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                http://aws.amazon.com/AWSECommerceService#PurchaseURL",
    "positive": true
  }]],
  "preconditions": [[{
    "id": "cart",
    "label": {"en-gb": "A shopping cart"},
    "pattern": "?C
                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                http://aws.amazon.com/AWSECommerceService#ShoppingCart",
    "positive": true
  }]],
  "code": "URL of the code",
  "definition": {
    "buildingblocks: [
      { // Form
        "id": "form1", // unique for the container screen
        "uri": "http://purl.oclc.org/fast/ontology/gadget#Form670238"
      },
      { // Operators
        "id": "op1",
        "uri": "http://purl.oclc.org/fast/ontology/gadget#Operator213487"
      },
      {  // Backend services
        "id": "bs1",
        "uri": "http://purl.oclc.org/fast/ontology/gadget#Backendservice38399"
      },
      {...}
    ]
    "pipes": [
      {
        "from": {
          "buildingblock": "bs1",
          "condition": "cA"
        },
        "to": {
          "buildingblock": "op1",
          "condition": "cA", // can be the same id, unique for the building block
          "action": "filter"
        }
      },
    "triggers": [
      {
        "from": {
          "buildingblock": "form1",
          "name": "refresh"
        },
        "to": {
          "buildingblock": "bs1",
          "action": "list"
        }
      }
    ]
  }
}
\end{verbatim}

\clearpage
\section*{Appendix E. Screen Component JSON Structure}
\addcontentsline{toc}{section}{Appendix E. Screen Component JSON Structure}
